Notes on Design and Implementation of Anglican

### 2015-08-19 paper/paper.tex
first finished draft

>>>>

A Clojure namespace that includes a definition of an Anglican
program imports (`requires') two essential namespaces:
\texttt{anglican.emit} and \texttt{anglican.runtime}. The former
provides macros for defining Anglican programs
(\texttt{defquery}, \texttt{query}) and functions
(\texttt{defm}, \texttt{fm}, \texttt{mem}), as well as Anglican
bootstrap definitions that must be included with every program
--- first of all, CPS implementations of higher-order functions.
\texttt{anglican.emit} can be viewed as the Anglican
\textit{compiler tool}, which helps transform Anglican code into
Clojure before any inference is performed.

\texttt{anglican.runtime} is the Anglican runtime library. For
convenience, it exposes common mathematical functions
(\texttt{abs}, \texttt{floor}, \texttt{sin}, \texttt{log},
\texttt{exp}, etc.), but most importantly, it provides
definitions of common distributions. Each distribution object
implements the \texttt{anglican.runtime/distribution} protocol,
with two methods: \texttt{sample} and \texttt{observe}. The
names of the methods coincide with Anglican special forms
\texttt{sample} and \texttt{observe}, but they are by no means
the same. The \texttt{sample} \textit{method} returns a random sample,
the \texttt{observe} \textit{method} returns the log probability
of the value. The methods can be, and sometimes are called from
handlers of the corresponding checkpoints, but do not have to
be. For example, in LMH either the \texttt{sample} or
the \texttt{observe} \textit{method} is called for a
\texttt{sample} checkpoint, depending on whether the value is
drawn or re-used.

Macro \texttt{defdist} should be used to define distributions.
\texttt{defdist} takes care of defining a separate record type
for every distribution so that multimethods can be dispatched on
distribution types when needed, e.g. for custom proposals.

Similar to distributions, \textit{random processes}, related to
so called `exchangeable random procedures', are defined using
macro \texttt{defproc} and implement the
\texttt{anglican.runtime/random-process} protocol. The protocol
has twho methods --- \texttt{produce}, which returns a
distribution object, and \texttt{absorb}, which returns a new
random process object obtained by absorbing a value drawn from
the process. While distributions are random and non-functional,
random processes are deterministic and functional, hence
\texttt{produce} and \texttt{absorb} are called directly and do
not have corresponding special forms in Anglican.
